---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(leaflet.circlesector)

# Example data
data("fans_tracks")
data("fans_mean")
data("fans_water")
data("outlier_segments")
data("light_segments_artificial")
data("light_segments_natural")

# Display helper
hr <- . %>%
  head() %>%
  reactable::reactable(
      sortable = TRUE, 
      filterable = TRUE, 
      searchable = TRUE, 
      defaultColDef = reactable::colDef(html = TRUE))
```

## Example data
The packaged example data is a subset from the package 
[wastdr](https://dbca-wa.github.io/wastdr/),
which itself is not included as a dependency to keep this package streamlined and small.

As a caveat, the data should be considered as synthetic and no real analysis
should be conducted on it.

As marine turtle hatchlings emerge from a nest and scamper towards what they believe
is the fastest way into the ocean, they form a fan of tracks in the sand.
The fan can be approximated through a circle segment from the leftmost to the rightmost
track's compass bearing. The number of tracks in a fan can sometimes be hard to precisely 
count, but can be estimated with a lower and upper estimate (e.g. ca 50 tracks,
min 30 max 70 tracks).

The raw data is documented at `help(fans)`.

Some hatchlings get the direction to water completely wrong and wander off on their
own, forming an outlier track. See also `help(outliers)`.
Some outliers are from nests not included in `fans`.

The biggest orientation help for hatchlings is the dim starlight over the ocean,
and the dark sand dune behind them. Since hatchlings wander towards the lightest
horizon, their orientation is particularly vulnerable to artificial light sources.
These are recorded, if known, as bearings from the nest location.
See also `help(lights)`.
Some light sources are from nests not included in `fans`.

For simplicity, the data is transmuted to have only the required columns
in the default naming scheme. This saves us from specifying the column names
when calling `addCircleSectorMinMax` and `addCircleSectorMid`.

The processing steps are kept in `inst/generate_data.R`.

The data we show here are:

* `fans_tracks`: Hatchling tracks form a main fan originating from the nest and spreading out.
* `fans_mean`: The mean direction of the main fan. This is where most hatchlings went.
* `fans_water`: The direction to water. This is where they are meant to go.
* `outlier_segments`: Outlier tracks. These folks got it completely wrong.
* `light_segments_natural`: Known natural light sources.
* `light_segments_artificial`: Known artificial light sources. These tend to disorient hatchlings.

Example data for `addCircleSectorMinMax`:

```{r}
fans_tracks %>% hr()
```

Example data for `addCircleSectorMid`:

```{r}
fans_mean %>% hr()
```

## Visualisation

At the location of each nests, we show:

* A white circle indicating the nest with radius 5m.
* The main turtle hatchling tracks fan as a blue circle segment from 
  `start_angle` to `end_angle` with radius 10m.
* A thin black segment pointing towards the water.
* A thin blue segment pointing towards the mean bearing.
* A thin red segment for each outlier.
* A thin yellow segment pointing towards each natural light source.
* A thin orange segment pointing towards each artificial light source.

```{r}
leaflet_basemap(l_height = 500, l_width = 700) %>%
  # clearBounds releases setView from leaflet_basemap
  # so that addCircleSectorMid/MinMax can expandLimits
  leaflet::clearBounds() %>%
  leaflet::addCircles(
    data = fans_tracks,
    lat = ~lat,
    lng = ~lon,
    color = "white",
    weight = 2,
    radius = 5
  ) %>%
  addCircleSectorMid(data = light_segments_artificial) %>%
  addCircleSectorMid(data = light_segments_natural) %>%
  addCircleSectorMid(data = fans_mean) %>%
  addCircleSectorMid(data = fans_water) %>%
  addCircleSectorMid(data = outlier_segments) %>%
  addCircleSectorMinMax(data = fans_tracks)
```
